GSE158422: DNA methylation in Lung squamous cell carcinoma - gene methylation enrichment analysis - sensitivity analysis

### 2023-12-20 figs/GSE158422_sensitivity.Rmd
use half of genes for sets
what about probe mean rank - do it all again in docker

CORES=24

@@

genesets <- gmt_import("https://ziemann-lab.net/public/gmea_prototype/ReactomePathways.gmt")

@@

```
## Run limma

@@

@@

tic()
dmagg <- agg(dm,cores=32)
m <- dmagg$gme_res_df[,"mean",drop=FALSE]
lamres <- runmitch(m=m,genesets=gs_symbols,cores=32)
toc()

@@

@@

SAMPLESIZES=c(2,3,6,10,16,22,30)
ssres <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100)
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    design2 <- design[mysample,]
    design2 <- design2[,colSums(design2)>0]
    mval2 <- mval[,mysample]
    dm2 <- runlimma(mval=mval2,design=design2,myann=myann)
    dmagg2 <- agg(dm2,cores=2)
    m2 <- dmagg2$gme_res_df[,"mean",drop=FALSE]
    lamres2 <- runmitch(m=m2,genesets=gs_symbols,cores=2)
    lamsig2 <- subset(lamres2,p.adjustANOVA<0.05)
    lamsigsets2 <- paste(rownames(lamsig2),as.character(lamsig2$s.dist > 0))
    TOT=length(lamsigsets2)
    TP=length(which(lamsigsets2 %in% lamsigsets))
    FP=length(which(!lamsigsets2 %in% lamsigsets))
    FN=length(which(!lamsigsets %in% lamsigsets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=5)
  do.call(rbind,dsres)
})

>>>>

CORES=16

@@

@@

## Run limma

@@

## LRM - limma rank mean mitch

```{r,lmrm1}

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #8.735s

dm$rank <-  rank(dm$t) - nrow(subset(dm,t<0))
mm <- merge(dm,gt,by.x=0,by.y="probe")
head(mm)
mma <-aggregate(mm$rank ~ gene,mm,mean)
rownames(mma) <- mma$gene
mma$gene = NULL
colnames(mma) <- "meanrank"
lrmres <- runmitch(m=mma,genesets=gs_symbols,cores=16)

lrmsig <- subset(lrmres,p.adjustANOVA<0.05)

top <- head(lrmsig,20)
top <- top[order(top$s.dist),]

hist(lrmsig$s.dist)

# distinguish up (TRUE) and down (FALSE)
lrmsigsets <- paste(rownames(lrmsig),as.character(lrmsig$s.dist > 0))
lrmsigsets2 <- gsub(" TRUE"," UP",lrmsigsets)
lrmsigsets2 <- gsub(" FALSE"," DN",lrmsigsets2)

@@

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #8.735s

tic()
dmagg <- agg(dm,cores=16)
m <- dmagg$gme_res_df[,"mean",drop=FALSE]
lamres <- runmitch(m=m,genesets=gs_symbols,cores=16)
toc()

@@

## LMRM - limma mean rank mitch

```{r,lmrm1}

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #8.735s

dm$rank <-  rank(dm$t) - nrow(subset(dm,t<0))
mm <- merge(dm,gt,by.x=0,by.y="probe")
head(mm)
mma <-aggregate(mm$rank ~ gene,mm,mean)
rownames(mma) <- mma$gene
mma$gene = NULL
colnames(mma) <- "meanrank"
lmrmres <- runmitch(m=mma,genesets=gs_symbols,cores=16)

lmrmsig <- subset(lmrmres,p.adjustANOVA<0.05)

top <- head(lmrmsig,20)
top <- top[order(top$s.dist),]

hist(lmrmsig$s.dist)

# distinguish up (TRUE) and down (FALSE)
lmrmsigsets <- paste(rownames(lmrmsig),as.character(lmrmsig$s.dist > 0))
lmrmsigsets2 <- gsub(" TRUE"," UP",lmrmsigsets)
lmrmsigsets2 <- gsub(" FALSE"," DN",lmrmsigsets2)

@@

blah()
SAMPLESIZES=c(2,3,6,10,16,22,30)
ssres <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100)
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    design2 <- design[mysample,]
    design2 <- design2[,colSums(design2)>0]
    mval2 <- mval[,mysample]
    dm2 <- runlimma(mval=mval2,design=design2,myann=myann)
    dmagg2 <- agg(dm2,cores=2)
    m2 <- dmagg2$gme_res_df[,"mean",drop=FALSE]
    lamres2 <- runmitch(m=m2,genesets=gs_symbols,cores=2)
    lamsig2 <- subset(lamres2,p.adjustANOVA<0.05)
    lamsigsets2 <- paste(rownames(lamsig2),as.character(lamsig2$s.dist > 0))
    TOT=length(lamsigsets2)
    TP=length(which(lamsigsets2 %in% lamsigsets))
    FP=length(which(!lamsigsets2 %in% lamsigsets))
    FN=length(which(!lamsigsets %in% lamsigsets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=5)
  do.call(rbind,dsres)
})
